---
output: 
  html_document: 
    keep_md: yes
---
---
title: "Peer-graded Assignment: Course Project 1"


```{r, echo=TRUE}
library(knitr)
library(plyr)  
library(ggplot2) 
opts_chunk$set(dev="png", fig.path="figure/")
```
#Loading and preprocessing the data
Load the data and show the first 6 rows of the data

```{r, echo=TRUE}
activity<-read.csv("activity.csv")
head(activity)
```

#What is mean total number of steps taken per day?
Ignoring the missing values in the dataset, here's the first 6 rows of total number of steps taken per day.

```{r, echo=TRUE}
day_step<-aggregate(steps~date,activity,sum)
head(day_step)
```

Histogram of the total number of steps taken each day

```{r, echo=TRUE}
hist(day_step$steps,main="Total steps in a day", col="blue",xlab="number of steps")
```



```{r, echo=TRUE}
m1<-mean(day_step$steps)
med1<-median(day_step$steps)
```

The mean is `r m1` and the median is `r med1`.

#What is the average daily activity pattern?
Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
interval_step<-aggregate(steps~interval,activity,mean)
plot(interval_step$interval,interval_step$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

```{r, echo=TRUE}
interval_max <- interval_step[which.max(interval_step$steps),1]
```
On average across all the days in the dataset, interval `r interval_max` contains the maximum number of steps


#Imputing missing values

```{r, echo=TRUE}
s1<-sum(is.na(activity$steps))
```
The total number of missing values in the dataset is `r s1`.
Some sample of data with missing values:
```{r, echo=TRUE}
head(activity)
```

Imputing the missing values with the mean of the relavant interval, here's a few rows of data with filled data
```{r, echo=TRUE}
interval_mean<-aggregate(steps~interval,activity,mean,na.rm=TRUE);
merge_data<-merge(activity,interval_mean,by="interval");
merge_data$steps.x[is.na(merge_data$steps.x)] <- merge_data$steps.y[is.na(merge_data$steps.x)];
merge_data<-merge_data[ ,1:3];
head(merge_data)
s<-sum(is.na(merge_data$steps.x))
```
The number of missing values of the filled data become `r s`.

Histogram of the total number of steps taken each day with filled data
```{r, echo=TRUE}
day_merge_data<-aggregate(steps.x~date,merge_data,sum)
hist(day_merge_data$steps.x,main="Total steps in a day with remove NA", col="blue",xlab="number of steps")
m2 <- mean(day_merge_data$steps.x)
med2<-median(day_merge_data$steps.x)
```
The mean of the imputed data is `r m2` and median is `r med2`. The mean is same as the first part but the median is higher than the first part.

#Are there differences in activity patterns between weekdays and weekends?
Creating new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
merge_data$date<-as.Date(merge_data$date, format = "%Y-%m-%d")
weekday<-weekdays(merge_data$date)
merge_data<-cbind(merge_data,weekday)
merge_data<- transform(merge_data, weekday = ifelse(weekday == "Saturday"|weekday == "Sunday", "weekend", "weekday"))
```

Some examples of the data with new variable:
```{r, echo=TRUE}
head(merge_data)
```

Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r, echo=TRUE}
average<-aggregate(steps.x ~ interval + weekday, merge_data, mean)
library(ggplot2)
ggplot(average, aes(interval, steps.x)) + geom_line() + facet_grid(weekday ~ .) + xlab("5 minutes interval") + ylab("Average of number of steps")
```
